# Deliverable 2

##### Team Member: Henry Bayly\|Jacqueline Lin\|Jiasong Huang\|Maysen Pagan\|Peiqi Lu\|Phil Ledoit

##### Lab session: C2

```{r setup, message=FALSE}
# Load visualization packages
library(tidyverse)
library(ggplot2)
library(hrbrthemes)
library(GGally)
#library(gridExtra)
library(paletteer)
```

We selected the **price** of every sold car as the response variable (Y) and the **mileage** of the car at the time of sale as the co-variate (X).

Our findings suggest that

> **there exists a negative correlation between the price of a car and its mileage at the time of sale.**

Typically, cars with lower mileage tend to be more valuable since they have undergone less wear and tear, potentially making them more dependable and long-lasting, albeit more costly. Conversely, cars with higher mileage usually command a lower price due to their greater likelihood of requiring frequent repairs and maintenance, leading to increased ownership expenses, hence cheaper than those with lower mileage.

### Data Loading

```{r,data loading}
bmw <- read.csv("BMWpricing.csv")
head(bmw)
```

### Original Scatter-plot

```{r,warning=FALSE}
#price to mileage
ggplot(bmw, aes(x = mileage, y = price,color=price)) + 
    geom_point(size = 0.2, alpha =0.2) + 
    geom_smooth(method = "lm", formula = y~x, size = 0.5, se = FALSE) +
    ylab("Price") + 
    xlab("Mileage") + 
    labs(color = "Price") + 
    labs(title = "Original Scatter-plot Price~Mileage")
```

### OLS Performance

```{r}
#std residuals
bmw.lm <- lm(price~mileage, data = bmw)
bmw_stdres <- rstandard(bmw.lm)
summary(bmw.lm)
```

#### output interpretation

Our model indicates that mileage is a significant factor in determining the price of a car, with a p-value of less than 2.2e-16 and an F-statistic of 980 on 1 and 4841 degrees of freedom. The linear regression model predicts that Price = 24670 - 0.06273\*Mileage, indicating that there is a negative correlation between price and mileage. This means that for every 1-mile increase in mileage, the price of the car decreases by 0.06273 dollars at the time of sale. Mileage has a p-value smaller than 2e-16, which indicates that it is a significant indicator of price. However, the model's adjusted R-squared value of 0.1682 suggests that only 16.82% of the variability in price can be explained by mileage. Mileage is a significant indicator of the price of the car, but it might not be the only indicator.

### **Standardized Residuals Plot**

```{r}
ggplot(bmw, aes(x = mileage, y = bmw_stdres,color=bmw_stdres)) + 
  geom_point(size = 0.2, alpha =0.2) + 
  labs(title = "BMW Prices",color = "std. RSS") +
  ylab("Standardized Residuals") + xlab("Mileage") + 
  geom_hline(yintercept=0, color = "blue") 
```

#### standardized residual plot interpretation

Based on the Standardized Residuals Plot, we can clearly see that the model assumptions of constant variance may not hold but rather decreases when Mileage increases and there are several outliers. So, our model is not a good model for predicting the price of a car at the time of sale.

### **Normal Q-Q Plot**

```{r}
 qqnorm(bmw$mileage, pch = 1, frame = FALSE)
 qqline(bmw$mileage, col = "steelblue", lwd = 2)
```

Based on the Normal Q-Q Plot, we can see that the data deviates for the straight line at the ends of the line, making the data not forming a rough staright line. This also indicates that our model is not a good model.
